---
title: "Replication of 'Physical activity and stress resilience: Considering those at risk for developing mental health problems' by Hegberg & Tone (2014, Mental Health and Physical Acitivty)"
author: "Jason Kaul (jbkaul@stanford.edu)"
date: "2022-10-09"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

## Introduction

I chose this particular study due to my passion for health and performance psychology, as well as to gain a deeper understanding into the implications of physical activity on stress-related mental health concerns in college students. As a current collegiate student-athlete, I have experienced the postive effects that physical activity (PA) can have on underlying stress levels. I was drawn to this study because of its utilization of questionnaires that measure 'self perceived resilience' and 'trait anxiety'; both measures that I beleive to be significant within the domain of mental health and performance. 

The main question this study aimed to address is simple, yet profound: "Does PA improve an individual's stress resilience, in turn, providing protection against stress-related mental health problems?". In the study, there were 222 participants that were all undergraduates enrolled at a Southern, urban univeristy and taking at least one psychology course. Researchers surveyed participants on their physical activity level (GPAQ-2), self-perceived resilience (CD-RISC 10), and trait anxiety (STAI-T) to test whether trait anxiety moderated the association between PA and self-percieved resilience. The researchers then conducted a multiple hierarchical linear regression analysis for their hypothesis testing. A potential challenge may be getting enough participants if we choose to narrow our population to only undergraduate students. A second challenge may be replicating the hierarchical linear regression that was performed in the original study.  

Repository: https://github.com/psych251/hegburg2014.git

Original Paper: https://github.com/psych251/hegburg2014/tree/main/original_paper

## Methods

### Power Analysis

Using G*Power 3.1, I conducted a power analyis that provided me a sample size of 161 participants with 99% power. The categories I selected within the power analysis were (t-tests, correlation point biserial model, a priori compute required sample size given a, power, and effect size, one tail, effect size of 0.3, a err prob of 0.05, and power of 0.99). I feel 160 particpants is a realistic number to anticipate taking my survey, however, there were a lot of different power analyses that led to different sample size totals. I think I will show a TA a screen shot of my power analysis to make sure my sample size makes sense.

### Planned Sample

"Participation was restricted to those who self-identified as proficient in the English language. Participants were asked to report their age, sex, and racial identity". These were the only demographic criteria listed in the article and I do not plan on asking any more demograhpic questions other than age, sex, and racial identity to keep my replication study as similar as possible to the original. 

### Materials

"Participants were 222 undergraduates enrolled at a Southern, urban university (see Table 1 for descriptive statistics). The sample's racial distribution was roughly consistent with that of the university student body and the local population, according to census data (U.S. Census Bureau, 2012). Participation was restricted to those who self-identified as proficient in the English language. Participants were recruited through an online research management system for students enrolled in psychology courses. They received detailed information about this IRB-approved study and about the voluntary nature of their participation. All data were collected using the Survey Monkey online questionnaire tool".

For methods, there isn't much needed since this is a correlational study done through online surveys. In terms of methods, I anticipate my replication study looking very similar to this. There will most likely not be as many participants (n=100-150), however, I am planning on having similar racial and age distributions. Not all participants will be university students and my replication study will be conducted using prolific. 

### Procedure	

The main procedure of this replication study will be creating the survey that my participants will take. In order to replicate the original study as closely as possible, I will use the same exact questionnaires that test the 3 main measures of interest: Physical Activity Level, Self-Percieved Resilience, and Trait Anxiety. 

Physical Activitiy Level:
"The Global Physical Activity Questionnaire Version 2 (GPAQ-2; World Health Organization [WHO], 2001) was used to assess
physical activity in a typical week across three contexts (i.e., activity at work, travel to and from places, recreational activities). The
GPAQ-2 includes 16 self-report items that measure time spent engaged in 1) vigorous- and moderate-intensity activity at work and during recreation, 2) walking and/or biking for transportation for at least 10 min continuously, and 3) sedentary behavior. The GPAQ-2 has been shown to be valid and reliable, as well as appropriate for use in varied cultural contexts (WHO, 2012). Recent research shows acceptable evidence of test-retest reliability, concurrent validity, criterion validity consistent with other measures of PA, and suggest that the GPAQ version 2 is suitable as a measure for monitoring PA (Bull, Maslin, & Armstrong, 2009; Herrmann, Heumann, Der Ananian, & Ainsworth, 2013)."

Self Percieved Resilience: 
"The Connor Davidson Resilience Scale 10 (CD-RISC 10; Campbell-Sills & Stein, 2007) was used to assess self-perceived
resilience in day-to-day life. The CD-RISC 10 is a unidimensional 10-item self-report measure designed to quantify resilience,
conceptualized as successful stress-coping ability. The measure yields an overall resilience score (0-40), with higher scores reflecting greater perceived resilience. The CDRISC 10 has been shown to have sound psychometric properties including reliability and validity across demographically diverse samples (Campbell-Sills & Stein, 2007; Coates, Phares, & Dedrick, 2013)."

Trait Anxiety: 
"The State-Trait Anxiety Inventory- Form X (STAI-T; Spielberger, Gorsuch, & Lushene, 1970) was used to assess trait anxiety. The
STAI-T is a 20-item self-report measure used to assess stable individual differences in anxiety proneness and scores reflect a general tendency to respond to the environment with anxiety (Spielberger, Gorsuch, Lushene, Vagg, & Jacobs, 1983).Item responses are summed to yield a total score ranging from 20 to 80, with higher scores indicating greater trait anxiety. The measure's author reported good test-retest reliability and inter-item consistency (Spielberger et al., 1970), which has since been replicated (Rule & Traver, 1983). The STAI-T has demonstrated good convergent and concurrent validity (Tilton, 2008), as well as convergent validity and strong psychometric properties across ethnic groups and gender (Novy, Nelson, Goodwin, & Rowzee, 1993)."


### Analysis Plan

Data Cleaning:
"Prior to conducting analyses, data were inspected for errors, excessive missing cases, outliers, and normality of distribution in accordance with published approaches (Bulmer, 1979; Hoaglin & Iglewicz, 1987; Tabachnick & Fidell, 2007; WHO, 2001). Where needed, data were transformed to better approximate normality. The distribution of GPAQ PA was positively skewed (SkewnessGPAQ ¼ 1.63, SE ¼ .21); therefore, these data were square root transformed (SkewnessGPAQ_transformed ¼ .44, SE ¼ .21). All other variables met assumptions of regression analysis".

Hypothesis Testing:
"To test whether trait anxiety moderated the association between PA and self-perceived resilience, we conducted a multiple hierarchical linear regression analysis. First, all predictors were either dummy coded (gender, race) or converted to z-scores, and then centered to facilitate interpretation of the findings and to reduce collinearity. At Step 1 of the regression model, covariates–Gender, Age, Race–were regressed onto the outcome variable (total CD-RISC score, which indexed self-perceived resilience). Demographic variables were included as covariates based on results of preliminary analyses indicating significant associations with study variables. At Step two, mean-centered trait anxiety and square root transformed and centered PA METs were added to the model. Lastly, at Step three, the trait anxiety  PA interaction term was added to the model. To obtain this variable, the centered trait anxiety total was multiplied by the transformed and centered METs of physical activity."

I plan on using a very similar method of statistical analysis as the one described in the original study. The offical title of the analysis is a "multiple hierarchial linear regression analysis". Along with the descriptions in the original paper, I plan on using this link - (https://data.library.virginia.edu/hierarchical-linear-regression/) - to assist in the data analysis. 

The one key result in the model will be to "test whether trait anxiety moderated the association between PA and self-perceived resilience". Although there are other relationships that can be measured, this is the main measure the authors were looking into. 

**multiple hierarchical linear regression analysis**

### Differences from Original Study

In terms of the survey questions and data analysis, I intend for these to be done as closely related to the orignal study as possible. The only difference I forsee happening are the amount and diveristy of our participants (in terms of age, race, SES, etc). I am deciding whether it would be best to recruit Stanford students in Psych 1 to take my survey, or gather all participants using Prolific. Taking Psych 1 students may be the closest I can get to the original study, however, I think I could get more participants using Prolific. 

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
  Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
  Any differences from what was described as the original plan, or “none”.


## Results


### Data preparation

Data preparation following the analysis plan.
	
	
```{r include=F}
###Data Preparation

####Load Relevant Libraries and Functions

library(tidyverse) # for data munging
library(knitr) # for table formatting
library(haven) # import and export 'SPSS', 'Stata' and 'SAS' Files
library(readxl) # import excel files
library(ggplot2) # ggplot2 library
library(lubridate)

####Import data

raw_data <- read_excel("~/desktop/Physical Activity & Stress Resilience.xlsx")


#### Prepare data for analysis - create columns etc.
```


### Cleaning Data


```{r}
# present all the column names 
colnames(raw_data)
```

```{r}
# attempt to get rid of miscellaneous headers
clean_data = raw_data %>%
  select(-StartDate, -EndDate, -Status, -IPAddress, -RecipientLastName, -RecipientFirstName, -RecipientEmail, -ExternalReference, -LocationLatitude,   -LocationLongitude, -DistributionChannel, -UserLanguage, -workerId, -assignmentId, -hitId, -ResponseId, -Prolific_ID) %>%
  mutate(Participant = row_number()) %>%
  relocate(Participant, demo_age, demo_gender, demo_race)

view(clean_data)
```


### Data exclusion / filtering


```{r}
# scoring of the GPAQ-2
GPAQ2_data = clean_data %>%
  select(starts_with("GPAQ2")) %>%
  mutate(across(c(GPAQ2_P2a, GPAQ2_P2b, GPAQ2_P4a, GPAQ2_P4b, GPAQ2_P6a,GPAQ2_P6b, GPAQ2_P8a, GPAQ2_P8b, GPAQ2_P10a, GPAQ2_P10b, GPAQ2_P11),               ~ifelse(str_detect(.,":"),as.numeric(hm(.))/60, as.numeric(.))))

GPAQ2_data <- replace(GPAQ2_data, is.na(GPAQ2_data), 0)              # replace NA values for 0
GPAQ2_data <- GPAQ2_data[-1,]                                        # get rid of first row containing survey questions

vigorous_work <- GPAQ2_data$GPAQ2_P2a * GPAQ2_data$GPAQ2_P2b * 6     # new variables to calculate MET levels 
moderate_work <- GPAQ2_data$GPAQ2_P4a * GPAQ2_data$GPAQ2_P4b * 3
vigorous_sport <- GPAQ2_data$GPAQ2_P8a * GPAQ2_data$GPAQ2_P8b * 9
moderate_sport <- GPAQ2_data$GPAQ2_P10a * GPAQ2_data$GPAQ2_P10b * 4.5

GPAQ2_data = GPAQ2_data %>%
  mutate(Total_MET = vigorous_work + moderate_work + vigorous_sport + moderate_sport)
  
head(GPAQ2_data)
```

```{r}
# histogram for GPAQ-2 responses
ggplot(GPAQ2_data, aes(x = GPAQ2_Total_MET)) + 
  geom_histogram(bins = 15, color = "black", fill = "azure") +
  labs(title = "GPAQ2 Physical Acitivity Level",
       x = "Physicial Activity Score",
       y = "Frequency")
```


```{r}
# scoring of the CD-RISC-10
RISC_data = clean_data %>%
  select(RISC_q_1, RISC_q_2, RISC_q_3, RISC_q_4, RISC_q_5, RISC_q_6, RISC_q_7, RISC_q_8, RISC_q_9, RISC_q_10) %>%
  mutate_at(vars(1:10), ~as.numeric(recode(.,
    "Not true at all" = "0", "Rarely true" = "1", "Sometimes true" = "2", "Often true" = "3", "True nearly all the time" = "4"))) %>%
  mutate(Total_Score = rowSums(.))
  na.omit(RISC_data)
  
RISC_data <- RISC_data[-1,]     # get rid of first row containing survey questions

head(RISC_data)
```

```{r}
# histogram for CD-RISK_-10 responses
ggplot(RISC_data, aes(x = Total_Score)) + 
  geom_histogram(bins = 20, color = "black", fill = "cornflowerblue") + 
  labs(title = "CD RISK 10 Questionnaire",
       x = "Resiliency Score",
       y = "Frequency")
```

```{r}
# scoring of the State-Trait Anxiety Inventory
STAIT_data = clean_data %>%
  select(starts_with("STAIT")) %>%
  mutate_at(vars(1:20), ~as.numeric(recode(.,
    "Almost Never" = "1", "Sometimes" = "2", "Often" = "3", "Almost Always" = "4"))) %>%
  mutate(`STAIT_q_1` = 4 - `STAIT_q_1`,
        `STAIT_q_3` = 4 - `STAIT_q_3`,
        `STAIT_q_6` = 4 - `STAIT_q_6`,
        `STAIT_q_7` = 4 - `STAIT_q_7`, 
        `STAIT_q_10` = 4 - `STAIT_q_10`,
        `STAIT_q_13` = 4 - `STAIT_q_13`,
        `STAIT_q_14` = 4 - `STAIT_q_14`,
        `STAIT_q_16` = 4 - `STAIT_q_16`,
        `STAIT_q_19` = 4 - `STAIT_q_19`,) %>%
  mutate(Total_Score = rowSums(.))

STAIT_data <- STAIT_data[-1,]    # get rid of first row containing survey questions
  
head(STAIT_data)
```

```{r}
# histogram for State-Trait Anxiety Inventory responses
ggplot(STAIT_data, aes(x = Total_Score)) + 
  geom_histogram(bins = 20, color = "black", fill = "lightpink3") + 
  labs(title = "State Trait Anxiety Inventory",
       x = "Trait Anxiety Score",
       y = "Frequency")
```

```{r}
# creating a new data frame that holds all key measures to be analyzed in the linear regression
RISK_Total_Score <- RISC_data$Total_Score
STAIT_Total_Score <- STAIT_data$Total_Score 
GPAQ2_Total_MET <- GPAQ2_data$Total_MET
Age <- as.numeric(clean_data$demo_age)
Gender <- clean_data$demo_gender
Race <- clean_data$demo_race

key_measures <- cbind(RISK_Total_Score, STAIT_Total_Score, GPAQ2_Total_MET, Age, Gender, Race)

key_measures <- key_measures[-1,] %>% as_tibble()   # get rid of first row containing survey questions

view(key_measures)
```

FINAL THOUGHTS -> I have started scoring for the different questionnaires, and completed the CD-RISC-10 scoring above. I am planning on completing the other two questionnaires very soon. My plan is to be in touch with the TA's throughout this next week as I work through what the linear regression will look like. I do feel like my qualtrics survey is ready to be uploaded to Prolific and hope to be given permission to start that process as soon as possible!


### Confirmatory analysis

The analyses as specified in the analysis plan.

```{r}

m1 <- lm(RISK_Total_Score ~ Gender + Age + Race, data = key_measures)                                           # Model 1
m2 <- lm(RISK_Total_Score ~ STAIT_Total_Score + GPAQ2_Total_MET + Gender + Age + Race, data = key_measures)     # Model 2
m3 <- lm(RISK_Total_Score ~ STAIT_Total_Score * GPAQ2_Total_MET + Gender + Age + Race, data = key_measures)     # Model 3

# anova1 <- anova(m0)  # to obtain Total SS
anova2 <- anova(m1, m2, m3)  # model comparison

summary(m1)
```


*Side-by-side graph with original graph is ideal here*

### Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
